---
title: "The Interview Attendance Problem"
author: "Jake Close and Peter Kaplan"
date: "11/21/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Context

The data pertains to the recruitment industry in India for the years 2014-2016 and deals with candidate interview attendance for various clients. When candidates don't appear for interviews, valuable time and resources are lost. If HR is able to predict which canidates may not appear, they could save money.



#importing libraries and reading data
```{r}
library(dummies)
library(class)
library(gmodels)
library(ggplot2)
library(dplyr)
library(C50)
interviews <- read.csv("Interview.csv")
```


#data cleaning

It appears the are extra columns read that are completely full of NA's. Lets drop these.


```{r}
interviews <- interviews[, 1:23]
```


Lets rename the variables to make them more readable
```{r}
names <- c("Date", "Client", "Industry", "Location", "Position", "Skillset", "Interview_Type", "ID", "Gender", "Candidate_Loc", "Job_Location", "Venue", "Native_Loc", "Permission", "Hope", "3_hour_call", "Alt_Number", "Resume_Printout", "Clarify_Venue", "Share_Letter", "Expected", "Attendance", "Marital_Status")
colnames(interviews) <- names
```


Next, lets explore which variables may need to be cleaned.


We check the intial structure of the dataset in order to do some data cleaning and exploration
```{r}
str(interviews)
```
From the intial structure, it appears that all of the variables are factor based. Lets explore these factors to make sure that there are no duplicates or NA

It appears that most of the columns have 1 NA. Lets see which if this is the last value?
```{r}
tail(interviews)
interviews <- interviews[1:1232, ]
```

#ID
For Classifation, we should drop Candidate ID since it is unique for each row.
```{r}
interviews <- interviews[, -8]
```



#industry
```{r}
levels(interviews$Industry)
```
It looks like there are three factors with IT in it. Lets combine these factors to one.

***needs word
```{r}
interviews[which(interviews$Location=='gurgaonr'),]$Location <- 'gurgaon'


#interviews[which(interviews$Industry =="IT Products and Services"), ]$Industry <- 'IT'

#interviews$Industry[interviews$Industry == "IT Products and Services"] <- "IT"
#interviews$Industry[interviews$Industry == "IT Services"] <- "IT"
levels(interviews$Industry)

barplot(table(interviews$Industry))
```

#Location
```{r}
levels(interviews$Location)
```
It looks like Chennai is spelt differently for four different factors, lets combine these. Also, Gurgaon is split with an extra r. We can fix this by converting all locations to lower

```{r}
interviews$Location <- tolower(interviews$Location)

```
```{r}
interviews$Location <- trimws(interviews$Location)

interviews[which(interviews$Location=='gurgaonr'),]$Location <- 'gurgaon'
table(interviews$Location)
```

#Position
```{r}
levels(interviews$Position)

```


```{r}
barplot(table(interviews$Position))
```

#Gender
```{r}
barplot(table(interviews$Gender))
```
We see that the data is much more male dominated.

#
```{r}
par(mfrow=c(2,2))
barplot(table(interviews$Hope))
title("Hope Distribution")
barplot(table(interviews$Permission))
title("Permission Distribution")
barplot(table(interviews$Alt_Number),las=2)
title("Alt number Distribution")
barplot(table(interviews$Resume_Printout),las=2)
title("Resume Distribution")
barplot(table(interviews$Clarify_Venue),las=2)
title("Clarify_Venue Distribution")

```
We find that these variables have many similar factors that should be combined.



## Native Location
Refers to the candidate's hometown.
```{r}
table(interviews$Native_Loc)
```
The majority of locations are <25, with the major candidate locations coming from Chennai, Hyderbad, and Bangalore.

##Job Location

```{r}
table(interviews$Job_Location)

```

#Distance function
Can we come up with a distance to job variable?



#Interview Type
There are three different types:
 Walkin drives, Scheduled, and scheduled walkin
```{r}
interviews$Interview_Type <- trimws(interviews$Interview_Type)
interviews$Interview_Type <- tolower(interviews$Interview_Type)
(table((interviews$Interview_Type)))
```
It appears the scheduled walkin is described in three ways. Lets combine these.
```{r}

```


#Expected
Lets clean up expected

```{r}
interviews$Expected <- trimws(interviews$Expected)
interviews$Expected <- tolower(interviews$Expected)
table(interviews$Expected)
```
Lets convert 10:30 am and 11:00 am to Yes

**need to do.
```{r}

```


#Client
```{r}
table(interviews$Client)
```
It looks like we can combine  Hewitt, Aon Hewitt and Aon Hewitt Gurgoan. Also Charted Bank?


###Marital Status
```{r}
barplot(table(interviews$Marital_Status))
```



Lets explore our predictor variable, observed attendance:
Again, we find this is dirty, so lets convert to lowercase first and trim.


```{r}
interviews$Attendance <- trimws(interviews$Attendance)
interviews$Attendance <- tolower(interviews$Attendance)
round(prop.table(table(interviews$Attendance)) * 100, 1)

```
We find about a 1/3 to 2/3 split of people that have attended.





